---
title: 'Learning novel skills from iconic gestures: A developmental and evolutionary
  perspective'
subtitle: 'Supplementary material and analysis'
date: "19 8 2018"
output:
  html_document:
    code_folding: hide
    theme: paper
    toc: yes
  pdf_document:
    toc: yes
---



# Overview

Please note that the some values in this document might be different compared to the ones reported in the corresponding manuscript. This is a consequence of the way the models are fit. Differences are minor and never affect interpretation of the results.

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(brms)
library(ggthemes)
library(tidyverse)
library(knitr)

```

## Datafile

Variables:

* subject: unique subject id.
* group: 2-, 3-year-olds or ape species.
* condition: arbitrary or iconic.
* apparatus: apart (app1 in manuscript) and boxing (app2 in manuscript)
* solve: successful box opening.
* sum_part_act: sum of partial actions (not applicable for 3-year-olds).

Head of data file:

```{r load data}

data <- read_csv("../data/ticon_data.csv")

head(data)%>%
  kable()
```

## Table for box opening by group and condition.
```{r descriptives }
data %>%
  mutate(group = as.character(group),
         group = ifelse (group == "24mo" | group == "36mo", group, "ape"))%>%
  group_by(group,condition)%>%
  summarise(n = n(),
            sum_solve = sum(solve))%>%
  kable()
```

## Overview plot (Figure 2 in manuscript)
```{r plot overview}

plot <- data %>%
  mutate(group = as.character(group),
         group = ifelse (group == "24mo" | group == "36mo", group, "Ape"),
         solve = factor(solve))
  

ggplot(plot,aes(condition))+
  geom_bar(aes(fill=solve),width = 0.4)+
  facet_grid( ~ group)+
  labs(x="",y="Nr. of Participants")+
  xlab("Condition")+
  scale_fill_manual(name="Open",
                    labels=c("No", "Yes"), values=c("lightgrey","#21912e"))+
  theme_few()+
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))+
  scale_y_continuous(breaks = seq(0, 18, by = 2))
```

# Study 1: 2- and 3-year-old children

## Models for box opening including both age groups

* m1child (model 1): `` solve ~ group * condition + apparatus``
* m2child (model 2): `` solve ~ group + condition + apparatus``
* m3child (model 3): `` solve ~ group + apparatus``

```{r models study 1 children, results='hide'}

## subsetting the data
child_data <- data %>%
  filter(group == "24mo" | group == "36mo")

# model including interaction

m1child = brm(solve ~ 
          group*condition + apparatus,
        data= child_data,
        family = bernoulli(),
        control = list(adapt_delta = 0.90),
        sample_prior = F,
        save_all_pars = T,
        iter = 10000)

# model without interaction

m2child = brm(solve ~ 
          group + condition + apparatus,
        data= child_data,
        family = bernoulli(),
        control = list(adapt_delta = 0.90),
        sample_prior = F,
        save_all_pars = T,
        iter = 10000)

# model without condition 

m3child = brm(solve ~ 
          group + apparatus,
        data= child_data,
        family = bernoulli(),
        control = list(adapt_delta = 0.90),
        sample_prior = F,
        save_all_pars = T,
        iter = 10000)

```

### WAIC

```{r model comparison child waic}

child_waic <- brms::WAIC(m1child, m2child,m3child, compare = F)

child_weights <- model_weights(m1child, m2child,m3child, weights = "waic")

data_frame(
  model = c("model 1","model 2","model 3"),
  WAIC = c(child_waic$m1child$estimates[3,1],child_waic$m2child$estimates[3,1],child_waic$m3child$estimates[3,1]),
  SE = c(child_waic$m1child$estimates[3,2],child_waic$m2child$estimates[3,2],child_waic$m3child$estimates[3,2]),
  weight = c(child_weights[1],child_weights[2],child_weights[3]))%>%
  kable(digits = 2)


```

### Bayes Factors

```{r bayes factor children model 1 vs model 2, results='hide'}

bf_child_12 <- bayes_factor(m1child, m2child, log = FALSE, maxiter = 5000)
bf_child_13 <- bayes_factor(m1child, m3child, log = FALSE, maxiter = 5000)
bf_child_23 <- bayes_factor(m2child, m3child, log = FALSE, maxiter = 5000)

```

```{r}
data_frame(
  comparison = c("model 1 vs. 2","model 1 vs. 3","model 2 vs. 3"),
  BF_10 = c(bf_child_12$bf,bf_child_13$bf,bf_child_23$bf))%>%
  kable(digits = 2)
```


### Summary for model including interaction (model 1)
```{r}
fixef(m1child)%>%
  kable(digits = 2)
```

#### Density plots

```{r plot for child model 1}

bayesplot::mcmc_areas(as.array(m1child), pars = c("b_Intercept", "b_group36mo", "b_conditioniconic", "b_group36mo:conditioniconic", "b_apparatusboxing"), prob = 0.95)+
  geom_vline(xintercept = 0, lty = 2, col = "black")+
  theme_few()

```

### Summary for model with only main effects (model 2)
```{r}
fixef(m2child)%>%
  kable(digits = 2)
```

#### Density plots

```{r plot for child model 2}

bayesplot::mcmc_areas(as.array(m2child), pars = c("b_Intercept", "b_group36mo", "b_conditioniconic", "b_apparatusboxing"), prob = 0.95)+
  geom_vline(xintercept = 0, lty = 2, col = "black")+
  theme_few()

```

### Figure S1: Comparing posterior predictions based on model 1 and 2

The plot below is based on samples drawn from the posterior distributions for model 1 (including interaction) and model 2 (only main effects).

Black triangles are expected proportions based on model 1, grey stars are based on model 2. Error bars are 95% intervals of extracted samples. Colored dots are condition means based on the data.
```{r plot posterior predictions}
library(rethinking)

post_m1 <- posterior_samples(m1child)%>%
  mutate(b3i = logistic(b_Intercept + b_apparatusboxing/2 +b_group36mo + b_conditioniconic + `b_group36mo:conditioniconic`),
         b3a = logistic(b_Intercept + b_apparatusboxing/2+ b_group36mo),
         b2i = logistic(b_Intercept + b_apparatusboxing/2+ b_conditioniconic),
         b2a = logistic(b_Intercept+ b_apparatusboxing/2))%>%
  select(b3i,b3a,b2i,b2a)%>%
  gather(pred,prop)%>%
  group_by(pred)%>%
  summarise(mean = quantile(prop,c(0.025,0.5,0.975)[2]),
            lci = quantile(prop,c(0.025,0.5,0.975)[1]),
            uci = quantile(prop,c(0.025,0.5,0.975)[3]))%>%
  ungroup()%>%
  mutate(group = c("24mo","24mo","36mo","36mo"),
         condition = c(0,1,0,1))%>%
  select(-pred)


post_m2 <- posterior_samples(m2child)%>%
  mutate(b3i = logistic(b_Intercept + b_apparatusboxing/2+b_group36mo + b_conditioniconic),
         b3a = logistic(b_Intercept + b_apparatusboxing/2+ b_group36mo),
         b2i = logistic(b_Intercept + b_apparatusboxing/2+ b_conditioniconic),
         b2a = logistic(b_Intercept+ b_apparatusboxing/2))%>%
  select(b3i,b3a,b2i,b2a)%>%
  gather(pred,prop)%>%
  group_by(pred)%>%
  summarise(mean = quantile(prop,c(0.025,0.5,0.975)[2]),
            lci = quantile(prop,c(0.025,0.5,0.975)[1]),
            uci = quantile(prop,c(0.025,0.5,0.975)[3]))%>%
  ungroup()%>%
  mutate(group = c("24mo","24mo","36mo","36mo"),
         condition = c(0,1,0,1))%>%
  select(-pred)


post_plot <- data %>%
  filter(group == "24mo" | group == "36mo")%>%
  group_by(group, condition)%>%
  summarise(mean = mean(solve))


ggplot()+
  geom_point(data = post_plot, aes(x = as.numeric(factor(condition)), y = mean, col=condition), size = 5)+
  geom_pointrange(data = post_m1, aes(x = as.numeric(condition)+0.8, y = mean, ymin = lci, ymax = uci), col = "black", shape = 2)+
  geom_pointrange(data = post_m2, aes(x = as.numeric(condition)+1.2, y = mean, ymin = lci, ymax = uci),col = "grey", shape = 8)+
  facet_grid( ~ group)+
  theme_few()+
  ylim(0,1)+
  xlim(0.5,2.5)+
  labs(x ="", y="Proportion Solved")+
  scale_color_solarized(name="Condition")+
  theme_few()+
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())
```


## Models for box opening by age group 


```{r models for condition by age group, results='hide' }
# subsetting data
data_24 <- data %>%
  filter(group == "24mo")

# subsetting data
data_36 <- data %>%
  filter(group == "36mo")

# model with condition 24 mo
m1_24 = brm(solve ~ 
          condition + apparatus,
        data= data_24,
        family = bernoulli(),
        control = list(adapt_delta = 0.90),
        sample_prior = F,
        save_all_pars = T,
        iter = 10000)

# model without condition 24 mo
m2_24 = brm(solve ~ 
          apparatus,
        data= data_24,
        family = bernoulli(),
        control = list(adapt_delta = 0.90),
        sample_prior = F,
        save_all_pars = T,
        iter = 10000)

# model with condition 36 mo
m1_36 = brm(solve ~ 
          condition + apparatus,
        data= data_36,
        family = bernoulli(),
        control = list(adapt_delta = 0.90),
        sample_prior = F,
        save_all_pars = T,
        iter = 10000)

# model without condition 36 mo
m2_36 = brm(solve ~ 
          apparatus,
        data= data_36,
        family = bernoulli(),
        control = list(adapt_delta = 0.90),
        sample_prior = F,
        save_all_pars = T,
        iter = 10000)

```

### 2-year-olds

#### WAIC

```{r condition model 24mo waic}
waic_24mo <- brms::WAIC(m1_24,m2_24, compare = F)

weights_24mo <- model_weights(m1_24, m2_24, weights = "waic")

data_frame(
  model = c("model w/ condition","model w/o conditon"),
  WAIC = c(
    waic_24mo$m1_24$estimates[3,1],
    waic_24mo$m2_24$estimates[3,1]),
  SE = c(waic_24mo$m1_24$estimates[3,2],
         waic_24mo$m2_24$estimates[3,2]),
  weight = c(weights_24mo[1],
             weights_24mo[2]))%>%
  kable(digits = 2)




```

#### Bayes Factor

```{r bayes factor for conditon for model for 24mo, results='hide'}
bf_24mo <- bayes_factor(m1_24, m2_24, log = FALSE, maxiter = 5000)
```

```{r}
data_frame(
  comparison = c("model w condition vs. w/o condition"),
  BF_10 = c(bf_24mo$bf))%>%
  kable(digits = 2)
```


#### Summary for model including condition

```{r summary for 24mo condition model}
fixef(m1_24)%>%
  kable(digits = 2)
```

#### Density plots

```{r plot for 24mo condition model}

bayesplot::mcmc_areas(as.array(m1_24), pars = c("b_Intercept","b_conditioniconic", "b_apparatusboxing"), prob = 0.95)+
  geom_vline(xintercept = 0, lty = 2, col = "black")+
  theme_few()

```

### 3-year-olds

#### WAIC

```{r condition model 36mo waic}
waic_36mo <- brms::WAIC(m1_36,m2_36, compare = F)

weights_36mo <- model_weights(m1_36, m2_36, weights = "waic")

data_frame(
  model = c("model w/ condition","model w/o conditon"),
  WAIC = c(
    waic_36mo$m1_36$estimates[3,1],
    waic_36mo$m2_36$estimates[3,1]),
  SE = c(waic_36mo$m1_36$estimates[3,2],
         waic_36mo$m2_36$estimates[3,2]),
  weight = c(weights_36mo[1],
             weights_36mo[2]))%>%
  kable(digits = 2)

```

#### Bayes Factor

```{r bayes factor for conditon for model for 36mo, results='hide'}
bf_36mo <- bayes_factor(m1_36, m2_36, log = FALSE, maxiter = 5000)

```

```{r}
data_frame(
  comparison = c("model w condition vs. w/o condition"),
  BF_10 = c(bf_36mo$bf))%>%
  kable(digits = 2)
```


#### Summary for model including condition

```{r summary for condition model 36mo}
fixef(m1_36)%>%
  kable(digits = 2)
```

#### Density plots

```{r plot for condition model 36mo}

bayesplot::mcmc_areas(as.array(m1_36), pars = c("b_Intercept","b_conditioniconic", "b_apparatusboxing"), prob = 0.95)+
  geom_vline(xintercept = 0, lty = 2, col = "black")+
  theme_few()

```

## Models for partial actions for 2-year-olds

### Figure S2: Plot for partial actions

Transparent dots show individual data points. Solid dots show means with 95% CI.

```{r plot partial actions 2-year-olds}

plot2 <- data %>%
  filter(group == "24mo")

library(gmodels)

plot3 <- data %>%
  filter(group == "24mo")%>%
  group_by(condition)%>%
  summarize(mean = mean(sum_part_act),
         lci = ci(sum_part_act)[2],
         uci = ci(sum_part_act)[3])
  
ggplot()+
  geom_jitter(data = plot2, aes(x = condition, y = sum_part_act, col = condition), height = 0, width = 0.2,alpha = 0.5)+
  geom_pointrange(data = plot3, aes(x = condition, y = mean, ymin = lci, ymax = uci, col = condition), size = 1)+
  labs(x="",y="Partial Actions")+
  scale_color_solarized(name = "Condition")+
  theme_few()+
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())


```


```{r model partial actions 24mo, results='hide'}

# model including condition

m1_24mopart = brm(sum_part_act ~ 
          condition + apparatus,
        data= data_24,
        family = poisson(),
        control = list(adapt_delta = 0.90),
        sample_prior = F,
        save_all_pars = T,
        iter = 10000)

# model excluding condition

m2_24mopart = brm(sum_part_act ~ 
          apparatus,
        data= data_24,
        family = poisson(),
        control = list(adapt_delta = 0.90),
        sample_prior = F,
        save_all_pars = T,
        iter = 10000)



```

### WAIC

```{r waic 24mo partial actions}
# model selection based on WAIC

m24_waic_part <- brms::WAIC(m1_24mopart, m2_24mopart, compare = F)

weights_24mo_part <- model_weights(m1_24mopart, m2_24mopart, weights = "waic")

data_frame(
  model = c("model w/ condition","model w/o conditon"),
  WAIC = c(
    m24_waic_part$m1_24mopart$estimates[3,1],
    m24_waic_part$m2_24mopart$estimates[3,1]),
  SE = c(m24_waic_part$m1_24mopart$estimates[3,2],
         m24_waic_part$m2_24mopart$estimates[3,2]),
  weight = c(weights_24mo_part[1],
             weights_24mo_part[2]))%>%
  kable(digits = 2)
```

### Bayes Factor

```{r bayes factor for condition 24mo partial actions, results='hide'}
bf_24mo_part <- bayes_factor(m1_24mopart,m2_24mopart, log = FALSE, maxiter = 5000)
```

```{r}
data_frame(
  comparison = c("model w condition vs. w/o condition"),
  BF_10 = c(bf_24mo_part$bf))%>%
  kable(digits = 2)
```


### Summary for model including condition

```{r summary for 24mo model partial actions}
fixef(m1_24mopart)%>%
  kable(digits = 2)
```

#### Density plots

```{r plot for 24mo model partial actions}

bayesplot::mcmc_areas(as.array(m1_24mopart), pars = c("b_Intercept","b_conditioniconic", "b_apparatusboxing"), prob = 0.95)+
  geom_vline(xintercept = 0, lty = 2, col = "black")+
  theme_few()

```

# Study 2: Apes 

## Models for box opening

```{r models study 2 apes, results='hide'}

## subsetting the data
ape_data <- data %>%
  filter(group != "24mo",
         group != "36mo")

## model including condition

m1ape = brm(solve ~ 
          condition + apparatus,
        data= ape_data,
        family = bernoulli(),
        control = list(adapt_delta = 0.90),
        sample_prior = F,
        save_all_pars = T,
        iter = 10000)

# model without condition 

m2ape = brm(solve ~ 
          apparatus,
        data= ape_data,
        family = bernoulli(),
        control = list(adapt_delta = 0.90),
        sample_prior = F,
        save_all_pars = T,
        iter = 10000)
```

### WAIC

```{r model comparison ape solve}

ape_waic <- brms::WAIC(m1ape,m2ape, compare = F)

ape_weights <- model_weights(m1ape, m2ape, weights = "waic")

data_frame(
  model = c("model w/ condition","model w/o conditon"),
  WAIC = c(
    ape_waic$m1ape$estimates[3,1],
    ape_waic$m2ape$estimates[3,1]),
  SE = c(ape_waic$m1ape$estimates[3,2],
         ape_waic$m2ape$estimates[3,2]),
  weight = c(ape_weights[1],
             ape_weights[2]))%>%
  kable(digits = 2)
```

### Bayes Factor

```{r bayes factor for condition ape model, results='hide'}

bf_ape <- bayes_factor(m1ape,m2ape, log = FALSE, maxiter = 5000)
```

```{r}
data_frame(
  comparison = c("model w condition vs. w/o condition"),
  BF_10 = c(bf_ape$bf))%>%
  kable(digits = 2)
```


### Summary for model including condition

```{r summary for ape model}

fixef(m1ape)%>%
  kable(digits = 2)
```

#### Density plots

```{r plot for ape model}

bayesplot::mcmc_areas(as.array(m1ape), pars = c("b_Intercept","b_conditioniconic", "b_apparatusboxing"), prob = 0.95)+
  geom_vline(xintercept = 0, lty = 2, col = "black")+
  theme_few()

```

## Model for partial actions

### Figure S3:Plot for partial actions

Transparent dots show individual data points. Solid dots show means with 95% CI.

```{r plot partial actions apes}

plot4 <- data %>%
  filter(group != "24mo",
         group != "36mo")

library(gmodels)

plot5 <- data %>%
  filter(group != "24mo",
         group != "36mo")%>%
  group_by(condition)%>%
  summarize(mean = mean(sum_part_act),
         lci = ci(sum_part_act)[2],
         uci = ci(sum_part_act)[3])
  
ggplot()+
  geom_jitter(data = plot4, aes(x = condition, y = sum_part_act, col = condition), height = 0, width = 0.2,alpha = 0.5)+
  geom_pointrange(data = plot5, aes(x = condition, y = mean, ymin = lci, ymax = uci, col = condition), size = 1)+
  labs(x="",y="Partial Actions")+
  scale_color_solarized(name = "Condition")+
  theme_few()+
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())

```


```{r model partial actions apes, results='hide'}

## model including condition 

m1apepart = brm(sum_part_act ~ 
          condition + apparatus,
        data= ape_data,
        family = poisson(),
        control = list(adapt_delta = 0.90),
        sample_prior = F,
        save_all_pars = T,
        iter = 10000)


# model without condition

m2apepart = brm(sum_part_act ~ 
          apparatus,
        data= ape_data,
        family = poisson(),
        control = list(adapt_delta = 0.90),
        sample_prior = F,
        save_all_pars = T,
        iter = 10000)

```

### WAIC

```{r waic ape partial actions}
ape_waic_part <- brms::WAIC(m1apepart,m2apepart, compare = F)

ape_weights_part <- model_weights(m1apepart, m1apepart, weights = "waic")

data_frame(
  model = c("model w/ condition","model w/o conditon"),
  WAIC = c(
    ape_waic_part$m1apepart$estimates[3,1],
    ape_waic_part$m2apepart$estimates[3,1]),
  SE = c(ape_waic_part$m1apepart$estimates[3,2],
         ape_waic_part$m2apepart$estimates[3,2]),
  weight = c(ape_weights_part[1],
             ape_weights_part[2]))%>%
  kable(digits = 2)
```

### Bayes Factor

```{r bayes factor ape partial actions, results='hide'}
bf_ape_part <- bayes_factor(m1apepart,m2apepart, log = FALSE, maxiter = 5000)
```

```{r}
data_frame(
  comparison = c("model w condition vs. w/o condition"),
  BF_10 = c(bf_ape_part$bf))%>%
  kable(digits = 2)
```

### Summary for model including condition

```{r summary for ape model partial actions}

fixef(m1apepart)%>%
  kable(digits = 2)

```

#### Density plots

```{r plot for ape model partial actions}

bayesplot::mcmc_areas(as.array(m1apepart), pars = c("b_Intercept","b_conditioniconic", "b_apparatusboxing"), prob = 0.95)+
  geom_vline(xintercept = 0, lty = 2, col = "black")+
  theme_few()

```


